### Note!!! This code is by: Juan Lopez-Martin, Justin H. Phillips, and Andrew Gelman, Multilevel Regression and Poststratification Case Studies, https://bookdown.org/jl5522/MRP-case-studies/downloading-and-processing-data.html

# Downloading and Processing Data

```{r packages-3, message=FALSE, echo = FALSE}
library(data.table)
library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(kableExtra)
library(ggplot2)
```

## CCES {#CCES}

The 2018 CCES raw survey data can be downloaded from the [CCES Dataverse](https://dataverse.harvard.edu)
via [this link](https://dataverse.harvard.edu/api/access/datafile/3588803?format=original&gbrecs=true);
by default the downloaded filename is called `cces18_common_vv.csv`.

Every MRP study requires some degree of data wrangling in order to make the factors in the survey of interest match the factors available in census data and other population-level surveys and census. Here we use the [R Tidyverse](https://www.tidyverse.org/) to process the survey data so that it aligns with the postratification table. Because initial recoding errors are fatal, it is important to check that each step of the recoding process produces expected results, either by viewing or summarizing the data.  Because the data is all tabular data, we use the utility function [`head`](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/head) to inspect the first few lines of a dataframe before and after each operation.

First, we examine the contents of the data as downloaded, looking only at those columns which provide the demographic-geographic information of interest. In this case, these are labeled as `inputstate`, `gender`, `birthyr`, `race`, and `educ`.

```{r results="asis", cache=FALSE}
cces_all <- read.csv("cces18_common_vv.csv")
```

```{r echo=FALSE, warning=FALSE, cache=FALSE}
kable(head(select(cces_all, inputstate, gender, birthyr, race, educ)), format='markdown')
```

As we have seen, it is crucial that the geography and demographics for the survery must match the geography and demographics in the poststratification table. If there is not a direct one-to-one relationship between the survey and the population data, the survey data must be recoded until a clean mapping exists. We write R functions to encapsulate the recoding steps.

We start considering the geographic information. Both the CCES survey and the US Census data use numeric [FIPS codes](https://transition.fcc.gov/oet/info/maps/census/fips/fips.txt) to record state information. We can use R factors to map FIPS codes to the standard two-letter state name abbreviations. Because both surveys use this encoding, we make this into a reusable function `recode_fips`.

```{r results="asis", cache=FALSE}
# Note that the FIPS codes include the district of Columbia and US territories which
# are not considered in this study, creating some gaps in the numbering system.
state_ab <- datasets::state.abb
state_fips <- c(1,2,4,5,6,8,9,10,12,13,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,
                31,32,33,34,35,36,37,38,39,40,41,42,44,45,46,47,48,49,50,51,53,54,55,56)
recode_fips <- function(column) {
    factor(column, levels = state_fips, labels = state_ab)
}
```

Secondly, we recode the demographics in order for them to be compatible with the American Community Survey data. In some cases this requires changing the levels of a factor (e.g. ethnicity) and in others we may need ot split a continous variable into different intervals (e.g. age). `clean_cces` uses the `recode_fips` function defined above to clean up the states. By default, the `clean_cces` functions drops rows where there is non-response in any of the considered factors or in the outcome variable; if this information is not missing at random, this introduces (more) bias into our survey.

```{r}
# Recode CCES
clean_cces <- function(df, remove_nas = TRUE){
  
  ## Abortion -- dichotomous (0 - Oppose / 1 - Support)
  df$abortion <- abs(df$CC18_321d-2)
  
  ## State -- factor
  df$state <- recode_fips(df$inputstate)
  
  ## Gender -- dichotomous (coded as -0.5 Female, +0.5 Male)
  df$male <- abs(df$gender-2)-0.5
  
  ## ethnicity -- factor
  df$eth <- factor(df$race,
                   levels = 1:8,
                   labels = c("White", "Black", "Hispanic", "Asian", "Native American", 
                              "Mixed", "Other", "Middle Eastern"))
  df$eth <- fct_collapse(df$eth, "Other" = c("Asian", "Other", "Middle Eastern", 
                                             "Mixed", "Native American"))
  
  ## Age -- cut into factor
  df$age <- 2018 - df$birthyr
  df$age <- cut(as.integer(df$age), breaks = c(0, 29, 39, 49, 59, 69, 120), 
                labels = c("18-29","30-39","40-49","50-59","60-69","70+"),
                ordered_result = TRUE)
  
  ## Education -- factor
  df$educ <- factor(as.integer(df$educ), 
                    levels = 1:6, 
                    labels = c("No HS", "HS", "Some college", "Associates", 
                               "4-Year College", "Post-grad"), ordered = TRUE)
  df$educ <- fct_collapse(df$educ, "Some college" = c("Some college", "Associates"))  
  
  # Filter out unnecessary columns and remove NAs
  df <- df %>% select(abortion, state, eth, male, age, educ) 
  if (remove_nas){
    df <- df %>% drop_na()
  }
  
  return(df)
}
```

## American Community Survey {#ACS}

We used the US American Community Survey to create a poststratification table. We will show two different (but equivalent) ways to do this. 

### Alternative 2: ACS PUMS {.unnumbered #ACS-PUMS}

Some researchers may prefer to access the 2018 5-year ACS data directly without using IPUMS, which makes the process less intuitive but also more reproducible. Additionally, this does not require creating an account, as the Public Use Microdata Sample (PUMS) from the ACS can be downloaded directly from the [data repository](https://www2.census.gov/programs-surveys/acs/data/pums/2018/5-Year/). The repository contains two .zip files for each state: one for individual-level variables and other for household-level variables. All the variables considered in our analysis are available in the individual-level files, but we will also download and process the household-level variable income to show how this could be done.

```{r, eval=FALSE}
# We start downloading all the zip files using wget. If you are using Windows you can download
# a pre-built wget from http://gnuwin32.sourceforge.net/packages/wget.htm
dir.create("poststrat_data/")
system('wget -O poststrat_data -e robots=off -nd -A "csv_*.zip" -R "index.html","csv_hus.zip",
"csv_pus.zip" https://www2.census.gov/programs-surveys/acs/data/pums/2018/5-Year/')
```

If this does not work, you can also access the [data repository](https://www2.census.gov/programs-surveys/acs/data/pums/2018/5-Year/) and download the files directly from your browser.

Once the data is downloaded, we process the .zip files for each state and then merge them together. IPUMS integrates census data accross different surveys, which results in different naming conventions and levels in some of the variables with respect to the PUMS data directly downloaded from the ACS repository. Therefore, the preprocessing steps is slightly different from the code shown above, but as the underlying data is the same we obtain an identical poststratification table.

```{r, eval=FALSE}
list_states_abb <- datasets::state.abb
list_states_num <- rep(NA, length(list_states_abb))

list_of_poststrat_df <- list()

for(i in 1:length(list_states_num)){
  # Unzip and read household and person files for state i
  p_name <- paste0("postrat_data/csv_p", tolower(list_states_abb[i]),".zip")
  h_name <- paste0("postrat_data/csv_h", tolower(list_states_abb[i]),".zip")
  p_csv_name <- grep('\\.csv$', unzip(p_name, list=TRUE)$Name, ignore.case=TRUE, value=TRUE)
  temp_df_p_state <- fread(unzip(p_name, files = p_csv_name), header=TRUE, 
                           select=c("SERIALNO","ST","CIT","PWGTP","RAC1P","HISP","SEX",
                                    "AGEP","SCHL"))
  h_csv_name <- grep('\\.csv$', unzip(h_name, list=TRUE)$Name, ignore.case=TRUE, value=TRUE)
  temp_df_h_state <- fread(unzip(h_name, files = h_csv_name),
                         header=TRUE, select=c("SERIALNO","FINCP"))
  
  # Merge the individual and household level variables according to the serial number
  temp_df <- merge(temp_df_h_state, temp_df_p_state, by = "SERIALNO")

  # Update list of state numbers that will be used later
  list_states_num[i] <-  temp_df$ST[1]

  ## Filter by citizenship
  temp_df <- temp_df %>% filter(CIT!=5)

  ## State
  temp_df$state <- temp_df$ST

  ## Gender
  temp_df$male <- abs(temp_df$SEX-2)-0.5

  ## Tthnicity
  temp_df$RAC1P <- factor(temp_df$RAC1P, 
                          levels = 1:9, 
                          labels = c("White", "Black", "Native Indian", "Native Alaskan",
                                     "Native Indian or Alaskan", "Asian", "Pacific Islander", 
                                     "Other", "Mixed"))
  temp_df$eth <- fct_collapse(temp_df$RAC1P, "Native American" = c("Native Indian", 
                                                                   "Native Alaskan",
                                                                   "Native Indian or Alaskan"))
  temp_df$eth <- fct_collapse(temp_df$eth, "Other" = c("Asian", "Pacific Islander", "Other", 
                                                       "Native American", "Mixed"))
  levels(temp_df$eth) <- c(levels(temp_df$eth), "Hispanic")
  temp_df$eth[(temp_df$HISP!=1) & temp_df$eth=="White"] <- "Hispanic"

  ## Age
  temp_df$age <- cut(as.integer(temp_df$AGEP), breaks = c(0, 17, 29, 39, 49, 59, 69, 120),
                     labels = c("0-17", "18-29","30-39","40-49","50-59","60-69","70+"),
                     ordered_result = TRUE)
  # filter out underages
  temp_df <- filter(temp_df, age!="0-17")
  temp_df$age <- droplevels(temp_df$age)

  ## Income (not currently used)
  temp_df$income <- cut(as.integer(temp_df$FINCP), 
                        breaks = c(-Inf, 9999, 19999, 29999, 39999, 49999, 59999, 69999, 79999, 
                                   99999, 119999, 149999, 199999, 249999, 349999, 499999, Inf),
                        ordered_result = TRUE, 
                        labels = c("<$10,000", "$10,000 - $19,999", "$20,000 - $29,999", 
                                   "$30,000 - $39,999", "$40,000 - $49,999",
                                   "$50,000 - $59,999", "$60,000 - $69,999",
                                   "$70,000 - $79,999","$80,000 - $99,999",
                                   "$100,000 - $119,999", "$120,000 - $149,999",
                                   "$150,000 - $199,999","$200,000 - $249,999",
                                   "$250,000 - $349,999", "$350,000 - $499,999",
                                   ">$500,000"))
  temp_df$income <- fct_explicit_na(temp_df$income, "Prefer Not to Say")

  ## Education
  temp_df$educ <- cut(as.integer(temp_df$SCHL), breaks = c(0, 15, 17, 19, 20, 21, 24), 
                      ordered_result = TRUE,
                      labels = c("No HS", "HS", "Some college", "Associates", 
                                 "4-Year College", "Post-grad"))
  temp_df$educ <- fct_collapse(temp_df$educ, "Some college" = c("Some college", "Associates"))
  
  # Calculate the poststratification table
  temp_df <- temp_df %>% drop_na(state, eth, male, age, educ, PWGTP) %>% 
    select(state, eth, male, age, educ, PWGTP)
  
  ## We sum by the inidividual-level weight PWGTP
  list_of_poststrat_df[[i]] <- temp_df %>% 
    group_by(state, eth, male, age, educ, .drop = FALSE) %>% 
    summarise(n = sum(as.numeric(PWGTP)))

  print(paste0("Data from ", list_states_abb[i], " completed"))
}

# Join list of state-level poststratification files
poststrat_df <- rbindlist(list_of_poststrat_df)
# Clean up state names
poststrat_df$state <- recode_fips(state)
# Write as csv
write.csv(poststrat_df, "poststrat_df.csv", row.names = FALSE)
```








Some researchers may prefer to access the 2018 5-year ACS data directly without using IPUMS, which makes the process less intuitive but also more reproducible. Additionally, this does not require creating an account, as the Public Use Microdata Sample (PUMS) from the ACS can be downloaded directly from the [data repository](https://www2.census.gov/programs-surveys/acs/data/pums/2018/5-Year/). The repository contains two .zip files for each state: one for individual-level variables and other for household-level variables. There are also two files, `csv_hus.zip` and `csv_hus.zip`, which contain these variables for all states. We will start creating a folder called `poststrat_data` and downloading these two files:

```{r, eval=FALSE}
# If you are using Windows you can download a pre-built wget from 
# http://gnuwin32.sourceforge.net/packages/wget.htm
dir.create("poststrat_data/")
system('wget -O poststrat_data2 -e robots=off -nd -A "csv_hus.zip","csv_pus.zip"
       https://www2.census.gov/programs-surveys/acs/data/pums/2018/5-Year/')
```

If this does not work, you can also download the files directly from the data repository.

Once the data is downloaded, we unzip the files and merge them together. IPUMS integrates census data accross different surveys, which results in different naming conventions and levels in some of the variables with respect to the PUMS data directly downloaded from the ACS repository. Therefore, the preprocessing steps is slightly different from the code shown above, but as the underlying data is the same we obtain an identical poststratification table. All the variables considered in our analysis are available in the individual-level files, but we will also download and process the household-level variable income in order to show how this could be done.

```{r, eval=FALSE}
list_states_abb <- datasets::state.abb
list_states_num <- rep(NA, length(list_states_abb))

# Unzip and read household and person files
p_name <- paste0("poststrat_data/csv_pus.zip")
h_name <- paste0("poststrat_data/csv_hus.zip")

p_csv_name <- grep('\\.csv$', unzip(p_name, list=TRUE)$Name, ignore.case=TRUE, value=TRUE)
unzip(p_name, files = p_csv_name, exdir = "poststrat_data")
p_csv_name = paste0("poststrat_data/", p_csv_name)
temp_df_p <- rbindlist(lapply(p_csv_name, fread, header=TRUE, 
                              select=c("SERIALNO","ST","CIT","PWGTP","RAC1P","HISP","SEX",
                                       "AGEP","SCHL")))

h_csv_name <- grep('\\.csv$', unzip(h_name, list=TRUE)$Name, ignore.case=TRUE, value=TRUE)
unzip(h_name, files = h_csv_name, exdir = "poststrat_data")
h_csv_name = paste0("poststrat_data/", h_csv_name)
temp_df_h <- rbindlist(lapply(h_csv_name, fread, header=TRUE, select=c("SERIALNO","FINCP")))

# Merge the individual and household level variables according to the serial number
temp_df <- merge(temp_df_h, temp_df_p, by = "SERIALNO")

# Exclude associated ares that are not states based on FIPS codes 
temp_df <- temp_df %>% filter(ST %in% state_fips)
temp_df$ST  <- factor(temp_df$ST, levels = state_fips, labels = state_ab)

## Filter by citizenship
temp_df <- temp_df %>% filter(CIT!=5)

## State
temp_df$state <- temp_df$ST

## Gender
temp_df$male <- abs(temp_df$SEX-2)-0.5

## Ethnicity
temp_df$RAC1P <- factor(temp_df$RAC1P, 
                        levels = 1:9, 
                        labels = c("White", "Black", "Native Indian", "Native Alaskan",
                                   "Native Indian or Alaskan", "Asian", "Pacific Islander", 
                                   "Other", "Mixed"))
temp_df$eth <- fct_collapse(temp_df$RAC1P, "Native American" = c("Native Indian", 
                                                                 "Native Alaskan",
                                                                 "Native Indian or Alaskan"))
temp_df$eth <- fct_collapse(temp_df$eth, "Other" = c("Asian", "Pacific Islander", "Other", 
                                                     "Native American", "Mixed"))
levels(temp_df$eth) <- c(levels(temp_df$eth), "Hispanic")
temp_df$eth[(temp_df$HISP!=1) & temp_df$eth=="White"] <- "Hispanic"

## Age
temp_df$age <- cut(as.integer(temp_df$AGEP), breaks = c(0, 17, 29, 39, 49, 59, 69, 120),
                   labels = c("0-17", "18-29","30-39","40-49","50-59","60-69","70+"),
                   ordered_result = TRUE)
# filter out underages
temp_df <- filter(temp_df, age!="0-17")
temp_df$age <- droplevels(temp_df$age)

## Income (not currently used)
temp_df$income <- cut(as.integer(temp_df$FINCP), 
                      breaks = c(-Inf, 9999, 19999, 29999, 39999, 49999, 59999, 69999, 79999, 
                                 99999, 119999, 149999, 199999, 249999, 349999, 499999, Inf),
                      ordered_result = TRUE, 
                      labels = c("<$10,000", "$10,000 - $19,999", "$20,000 - $29,999", 
                                 "$30,000 - $39,999", "$40,000 - $49,999",
                                 "$50,000 - $59,999", "$60,000 - $69,999",
                                 "$70,000 - $79,999","$80,000 - $99,999",
                                 "$100,000 - $119,999", "$120,000 - $149,999",
                                 "$150,000 - $199,999","$200,000 - $249,999",
                                 "$250,000 - $349,999", "$350,000 - $499,999",
                                 ">$500,000"))
temp_df$income <- fct_explicit_na(temp_df$income, "Prefer Not to Say")

## Education
temp_df$educ <- cut(as.integer(temp_df$SCHL), breaks = c(0, 15, 17, 19, 20, 21, 24), 
                    ordered_result = TRUE,
                    labels = c("No HS", "HS", "Some college", "Associates", 
                               "4-Year College", "Post-grad"))
temp_df$educ <- fct_collapse(temp_df$educ, "Some college" = c("Some college", "Associates"))

# Calculate the poststratification table
temp_df <- temp_df %>% drop_na(state, eth, male, age, educ, PWGTP) %>% 
  select(state, eth, male, age, educ, PWGTP)

## We sum by the inidividual-level weight PWGTP
poststrat_df <- temp_df %>% 
  group_by(state, eth, male, age, educ, .drop = FALSE) %>% 
  summarise(n = sum(as.numeric(PWGTP)))

# Write as csv
write.csv(poststrat_df, "poststrat_df.csv", row.names = FALSE)
```

